Use of asynchronous circuits and leakiness in modeling biological circuit of the Caulobacter Crescentus cell cycle

### 2017-04-05 asyncleakyn.tex
added to hold times in VLSI

\documentclass{article}
\usepackage[utf8]{inputenc}

@@

Memory elements are important in asynchronous circuits because they can hold the signal for some time until they receive input, usually from elements further down the circuit to signal that the current output has been acknowledged and processed already and does not need to hold anymore.  This can be implemented in a number of ways but one method is adding a capacitor to the output.  The equation of a capacitor is 
\[I(t)=C\frac{dV(t)}{dt}
\]
It can be thought of as being charged up by the voltage as a high voltage is applied from $V_{out}$, or discharges when $V_{out}$ is low.  When in an RC circuit, the capacitor charges and discharges exponentially.  This is a valid way of maintaining the output voltage at either high or low for short amounts of time. 
\begin{center}
\includegraphics{mosinv}
\end{center}

@@

\section{Dynamic Memory}
%some intro about memory?
%EE
We have seen in the previous section how short term memory elements can be implemented with capacitors.  We examine some other ideal memory elements we would like as part of an asynchronous circuit, such as the Muller C-element. 
The Muller C-element is useful in the design of asynchronous circuits.  If every signal in a circuit is connected with C-elements then it ensures that the circuit can run robustly and has the speed independence property.  Thus we would ideally want the elements we see in biological circuits to operate similarly to the C-element.  A common CMOS implementation of a C-element is shown with two inputs A and B, and an output C.  It operates similarly as the inverter example.  In this C-element example, when A and B are both high, the pull down network is activated, while there are open switches in the pull up network and the output is connected to the ground.  When both A and B are low, the opposite is true and the output is pulled to $V_{cc}$.  However when A and B are different signals ie A is high and B is low or when A is low and B is high, then there is no path through either the pull up or the pull down network and the previous value of the output is retained.  This state is called a 'hold'.
\begin{center}
\includegraphics{celement}
\end{center}
\begin{center}

@@

@@

In our model we wish to verify the set reset and hold property of each gene gate.  We use the exponential solution to the ODE equation.  Given the appropriate inputs we would like for all of the set and and reset transitions to take a shorter length of time than the temporary hold states.  More formally, we would like the worst case/longest set and reset transition times to be shorter than the worst case/shortest hold times.  First we define some of the parameters and variables that we use.  Define the input thresholds as:  $in_{min0}$ is the smallest acceptable input to represent an input '0' and $in_{max0}$ is the largest acceptable input to represent an input '1'.  Similarly, $in_{min1}$ is the smallest acceptable input to represent an input '1' and $in_{max1}$ is the largest acceptable input to represent an input '1'.  $fanout_0$ and $fanout_1$ represent the '0' and '1' input thresholds to the next components respectively.  There is an additional constraint where fanout thresholds $fanout_0$ and $fanout_1$ must be between $g(in_{max0})$ and $g(in_{min1})$ so that all legal input values provide legal outputs that feed into the next layer of components to be processed.

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

@@

Memory elements are important in asynchronous circuits because they can hold the signal for some time until they receive input, usually from elements further down the circuit to signal that the current output has been acknowledged and processed already and does not need to hold anymore.  This can be implemented in a number of ways but one method is adding a capacitor to the output.  The equation of a capacitor is 
\[I(t)=C\frac{dV(t)}{dt}
\]
It can be thought of as being charged up by the voltage as a high voltage is applied from $V_{out}$, or discharges when $V_{out}$ is low.  When in an RC circuit, the capacitor charges and discharges exponentially.  This is a valid way of maintaining the output voltage at either high or low for short amounts of time.  We work these t in detail in a later section. 
\begin{center}
\includegraphics{mosinv}
\end{center}

@@

\section{Dynamic Memory}
%some intro about memory?
%EE
We have seen in the previous section how short term memory elements can be implemented with capacitors.  We examine some other ideal memory elements we would like as part of an asynchronous circuit, such as the Muller C-element. 
The Muller C-element is useful in the design of asynchronous circuits.  If every signal in a circuit is connected with C-elements then it ensures that the circuit can run robustly and has the speed independence property.  Thus we would ideally want the elements we see in biological circuits to operate similarly to the C-element.  A common CMOS implementation of a C-element is shown with two inputs A and B, and an output C.  It operates similarly as the inverter example.  In this C-element example, when A and B are both high, the pull down network is activated, while there are open switches in the pull up network and the output is connected to the ground.  When both A and B are low, the opposite is true and the output is pulled to $V_{cc}$.  However when A and B are different signals ie A is high and B is low or when A is low and B is high, then there is no path through either the pull up or the pull down network and the previous value of the output is retained.  This state is called a 'hold'.  Note that this is different from a hold time, which is a synchronous digital circuit concept.  Hold time is the amount of time that a valid signal retains its value.  It is useful to know to ensure that the signal maintains its correct value to allow downstream gates to properly compute the output. This is also important to our system since we are using dynamic memory.
%added hold time
\begin{center}
\includegraphics{celement}
\end{center}
\begin{center}

@@

\end{center}
As stated previously, the equation for a capacitor is:

\[I(t)=C\frac{dV(t)}{dt}
When we analyze the CMOS inverter connected to a load capacitance, when the input is high, the p-MOS is effectively off and the n-MOS operating in linear region can be approximately modeled as a resistor.  Equating the current from the resistor and capacitor, we find that the equation for Vout is:
\[\frac{-V_{out}(t)}{R}=C\frac{dV_{out}(t)}{dt}
When the input is low, the n-MOS is effectively off and the p-MOS operating in linear region can be approximated as a resistor.  Equating the current from the resistor and capacitor, the equation for Vout is:
\[\frac{Vdd-V_{out}(t)}{R}=C\frac{dV_{out}(t)}{dt}
\includegraphics[width=\textwidth]{mosinactionRC}
These differential equations can be written in a more general form of 
\[\frac{dV(t)}{dt}=C -\lambda \cdot V(t)
Where $C$ and $\lambda$ are constants and the solution to this differential equation is an exponential of the form
\[V(t)=((V_0\cdot\lambda-C)e^{(-\lambda\cdot t)}+C)/\lambda
where $V_0$ is the initial value of V at time t=0
Next we define some timing terminology.   Define the input thresholds as:  $in_{min0}$ is the smallest acceptable input to represent an input '0' and $in_{max0}$ is the largest acceptable input to represent an input '1'.  Similarly, $in_{min1}$ is the smallest acceptable input to represent an input '1' and $in_{max1}$ is the largest acceptable input to represent an input '1'.   $out_{max0}$ and $out_{min1}$ corresponds to the output signal after putting the input signal at the appropriate input thresholds through the gate.  $fanout_0$ and $fanout_1$ represent the '0' and '1' input thresholds to the next components respectively. \newline \newline
Three values are of importance to us, the transition time between 0->1, the transition time between 1->0, and the time that the gate is in the hold state.  Using the inverter example, the transition time between 0->1 is when the output is initially a valid 0 (so the initial input is 1), at time t=0, the input changes to 0 which prompts the output to change to 1.  We assume an output low threshold of 0.8V ($out_{max0}$) and a fanout high threshold of 4V ($fanout_1$).  The transition time is the time it takes between the output from below the low threshold until the ouput reaches the high fanout threshold.
\includegraphics[width=\textwidth]{t0to1}
Rearranging the equation to find time elapsed:
\[\Delta t_{0->g1}= -log(V_{ss}-fanout_1)/ \lambda+ log(V_{ss}-V_0)/ \lambda
Where $V_{ss}$ is the steady state output of the gate after the output finishes transitioning, usually Vdd in this case. \newline \newline
Similarly, the transition time between 1->0 is when the output transitions from a 1 to a 0.  We assume an output high threshold of 4.5V($out_{min1}$) and a fanout low threshold of 1V ($fanout_0$). \newline
\includegraphics[width=\textwidth]{t1to0}
The time elapsed is:
\[\Delta t_{1->0}= -log(V_0-V_{ss})/ \lambda+ log(fanout_0-V_{ss})/ \lambda
Where $V_{ss}$ is the steady state output of the gate after the output finishes transitioning, usually 0V in this case. \newline \newline
The time a gate stays in the 'hold' state, is when the output switches and the charge on the capacitor discharges or accumulates but still has a valid 1 or 0 for some time.  So it is the time it takes between a high output to reach fanout high threshold as it is discharging, or a low output to reach fanout low threshold as it is charging. \newline
\includegraphics[width=\textwidth]{hold0to1}
\includegraphics[width=\textwidth]{hold1to0}
The corresponding elapsed time are:
\[\Delta t_{hold 0}= -log(V_{ss}-fanout_0)/ \lambda+ log(V_{ss}-V_0)/ \lambda
\[\Delta t_{hold 1}= -log(V_0-V_{ss})/ \lambda+ log(fanout_1-V_{ss})/ \lambda
Where $V_{ss}$ is usually Vdd in the first equation and 0V in the second one. Note that for an inverter the time for hold state is a lot less than the transitions from 0 to 1 or 1 to 0.\newline \newline

@@

In our model we wish to verify the set reset and hold property of each gene gate.  We use the exponential solution to the ODE equation.  Given the appropriate inputs we would like for all of the set and and reset transitions to take a shorter length of time than the temporary hold states.  More formally, we would like the worst case/longest set and reset transition times to be shorter than the worst case/shortest hold times.  First we define some of the parameters and variables that we use.  Define the input thresholds as the previous section:  $in_{min0}$ is the smallest acceptable input to represent an input '0' and $in_{max0}$ is the largest acceptable input to represent an input '1'.  Similarly, $in_{min1}$ is the smallest acceptable input to represent an input '1' and $in_{max1}$ is the largest acceptable input to represent an input '1'.  $fanout_0$ and $fanout_1$ represent the '0' and '1' input thresholds to the next components respectively.  There is an additional constraint where fanout thresholds $fanout_0$ and $fanout_1$ must be between $g(in_{max0})$ and $g(in_{min1})$ so that all legal input values provide legal outputs that feed into the next layer of components to be processed.
